---
title: "PCA Lab"
author: "DM"
date: "March 23, 2026"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##EXERCISE I

## Data Reading
```{r include = TRUE}
A = matrix(c(9,12,10,15,9,10,5,10,8,11,13,14,11,13,8,3,15,10), nrow = 6, byrow = TRUE)
Nframe = as.data.frame(A)
m1 = c("Alexis", "Bea", "Claude","Damien", "Emilie", "Francis")
m2 = c("Physics", "Maths", "English")
row.names(A) = m1
colnames(A) = m2
head(A, 4)
mean(A[,1])
# Covariance matrix
cov(A)*(5/6)
# Correlation matrix
cor(A)
```

## PCA Function


```{r}
library(FactoMineR)
library(factoextra)
# Unscaled PCA
#res.pca <- PCA(A, graph = FALSE, scale.unit = FALSE)
# Scaled PCA
res.pca <- PCA(A, graph = FALSE)
print(res.pca)
```

## Eigenvalues

```{r}
res.pca$eig
eig.val = res.pca$eig
eig.val
# or
eig.val <- get_eigenvalue(res.pca)
eig.val
```


## Results for Variables

```{r}
var <- get_pca_var(res.pca)
var
```

## Correlations between variables and components

```{r}
head(var$cor)
```

## Coordinates of variables

```{r}
head(var$coord)
```

## Quality of representation of variables

```{r}
head(var$cos2)
```

## Contributions of variables

```{r}
head(var$contrib,3)
```


## With quality of representation

```{r}
fviz_pca_var(res.pca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE)
```


```{r}
# Contributions des variables ? la premi?re composante 
fviz_contrib(res.pca, choice = "var", axes = 1, top = 10)
# Contributions of variables ? la deuxi?me composante 
fviz_contrib(res.pca, choice = "var", axes = 2, top = 10)
```

## Contributions

```{r}
# Mise en valeur des plus importantes 
fviz_pca_var(res.pca, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")
             )
```
## The Individuals

```{r}
res.desc <- dimdesc(res.pca, axes = c(1,2), proba = 0.05)
res.desc$Dim.2
ind <- get_pca_ind(res.pca)
ind
```
## Individuals (Dimension 1)

```{r}
res.desc$Dim.1

```
## Individuals: Coordinates

```{r}
head(ind$coord)
```
## Quality of representation

```{r}
# Qualit?
head(ind$cos2)
round(ind$cos2, 2)
```
## Contributions

```{r}
head(ind$contrib)
```

## Contributions on first two components

```{r include=TRUE}
fviz_contrib(res.pca, choice = "ind", axes = 1:2)
```
## Individuals plot

```{r include=TRUE}
fviz_pca_ind(res.pca, col.ind = "cos2",gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE)
```












